On the presentation of orthogonal groups over polynomial rings

### 2019-01-29 dl-paper.tex
Writing down Proposition 1

\subsection{Unstable Quillen K-groups}
In order to better understand $C(\Phi, R, I)$  we would like to continue the sequence~\eqref{eq:relative-St} to the left. 
Thus we need to work with unstable $\K_3$-functors a la Quillen (cf.~\cite{Pa89}).

Let $G \colon \catname{CRings} \to \catname{Groups}$ be a functor preserving binary pullbacks and 
 $P$ its subfunctor such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$.

Thanks to the functoriality of Quillen's plus-construction (see e.\,g.~\cite[Proposition~5.2.4]{Ro95}) one can define a functor 
$X_{G, P}(R) = BG(R)^+_{P(R)} \colon \catname{CRings} \to \catname{Top}$ and then define the corresponding K-group functors via $\K_{i}(G, P, R) = \pi_i(X_{G, P}(R))$.

\begin{prop}\label{characterization}
Let $S \colon \catname{CRings} \to \catname{Groups}$ be another group functor equipped with a natural transformation
 $\varphi \colon S \to P$ such that for all $R$ the group $S(R)$ is a universal central extension of $P(R)$. Then the following holds:
\begin{proplist}
 \item There is a natural isomorphism $\K_3(G, P, R) \cong \HH_3(S(R), \ZZ)$.
 \item There is an exact sequence natural in $(R, I)$:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & \HH_3(S(R), S(R/I), \ZZ).} \]
\end{proplist}
\end{prop}
\begin{proof}
 The first claim follows from the solution of Exercises~1.8--1.9 of \cite[Chapter~IV]{Kbook}. %provided one verifies that all the intermediate isomorphisms are natural in $R$.

@@

 Since $\varphi(G_i) \subseteq \Ker(G(p_i))$ and $G$ preserves pullbacks we conclude that $G_0 \cap G_1 \subseteq \Ker(\varphi) \colon S(D_{R, I}) \to G(D_{R, I}))$
  and hence $G_0 \cap G_1$ is a central subgroup of $S(D_{R, I})$.
 Now invoking~\cite[Proposition~6]{Lo78} we conclude that the extension $S(R, I) \to S(R)$ is a universal
  relative central extension of the coequalizer $\mathrm{coeq}(d_0, d_1)$.

@@

\begin{lemma} Let $\Phi$ be a reduced irreducible root system of type $\neq \rG_2$ and $k$ be arbitrary field.
Then there is a split exact sequence of abelian groups
\[ \xymatrix{ 0 \ar[r] & \K_2(\Phi, k) \ar[r] & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & H(\Phi, k) \ar[r] & 0,}\text{ where} \]
\[ H(\Phi, k) = \left\{\begin{array}{ll} \K_1^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ k^* & \text{otherwise.}  \end{array}\right. \]  \end{lemma}
\begin{proof} In the nonsymplectic case one can choose a long simple root $\alpha$ of $\Phi$ in such a way that there is a commutative diagram of abelian groups
\[\xymatrixcolsep{5pc}\xymatrix{k^* \ar[r]^-{\{ -, X \}_{\alpha}} \ar[rrd]_{\{-, X\}} & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & \K_2(\Phi, k(X)) \ar[d]^{\cong} \\
                                                                                      &                                 & \K_2^\mathrm{M}(k(X)).} \]
Clearly, the diagonal arrow is injective by Milnor's theorem and the vertical arrow is an isomorphism by Matsumoto's theorem~\cite[Theorem~5.10]{Ma69}.
The required assertion now follow from~\cite[Satz~3]{Hur77} and the fact that the left horizontal arrow is injective.

>>>>

\subsection{Unstable Quillen K-groups}
\begin{comment}
In order to better understand $C(\Phi, R, I)$  we would like to continue the sequence~\eqref{eq:relative-St} to the left. 
Thus we need to work with unstable $\K_3$-functors a la Quillen (cf.~\cite{Pa89}).
\end{comment}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following assumptions:
\begin{enumerate}
 \item\label{req:left-exact} $G$ is left exact;
 \item $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coqualizers of kernel pairs.
\end{enumerate}

Thanks to the functoriality of Quillen's plus-construction (see e.\,g.~\cite[Proposition~5.2.4]{Ro95}) one can define a functor 
$X_{G, P}(R) = BG(R)^+_{P(R)} \colon \catname{CRings} \to \catname{Top}$ and then define for $i \geq 1$ the corresponding K-group functors $\K_{i}(G, P, R)$ as the homotopy groups $\pi_i(X_{G, P}(R))$.

\begin{prop}\label{characterization}
 Then the following holds:
\begin{proplist}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item There is an exact sequence natural in $(R, I)$:
 \end{proplist}
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & \HH_3(S(R), S(R/I), \ZZ).} \]
\end{prop}
\begin{proof}
 The first claim follows from the solution of Exercises~1.8--1.9 of \cite[Chapter~IV]{Kbook}. %provided one verifies that all the intermediate isomorphisms are natural in $R$.

@@

 From~\eqref{req:left-exact} and $\varphi(G_i) \subseteq \Ker(G(p_i))$ we conclude that $G_0 \cap G_1 \subseteq \Ker(\varphi) \colon S(D_{R, I}) \to G(D_{R, I})$
  and hence $G_0 \cap G_1$ is a central subgroup of $S(D_{R, I})$.
 Now invoking~\cite[Proposition~6]{Lo78} we conclude that the extension $S(R, I) \to S(R)$ is a universal relative central extension of the coequalizer $\mathrm{coeq}(d_0, d_1)$, 
   which coincides with $S(R) \to S(R/I)$ by~\eqref{req:coeq}.

@@

\begin{prop} Let $\Phi$ be a reduced irreducible root system of type $\neq \rG_2$ and $k$ be arbitrary field.
Then there is a split exact sequence of abelian groups
\[ \xymatrix{ 0 \ar[r] & \K_2(\Phi, k) \ar[r] & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & H(\Phi, k) \ar[r] & 0,}\text{ where} \]
\[ H(\Phi, k) = \left\{\begin{array}{ll} \K_1^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ k^* & \text{otherwise.}  \end{array}\right. \]  \end{prop}
\begin{proof} In the nonsymplectic case one can choose a long simple root $\alpha$ of $\Phi$ in such a way that there is a commutative diagram of abelian groups
\[\xymatrixcolsep{5pc}\xymatrix{k^* \ar[r]^-{\{ -, X \}_{\alpha}} \ar[rrd]_{\{-, X\}} & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & \K_2(\Phi, k(X)) \ar[d]^{\cong} \\
                                                                                      &                                 & \K_2^\mathrm{M}(k(X)).} \]
Clearly, the diagonal arrow is injective by Milnor's theorem and the vertical arrow is an isomorphism by Matsumoto's theorem~\cite[Theorem~5.10]{Ma69}.
The required assertion now follow from~\cite[Satz~3]{Hur77} and the fact that the left horizontal arrow is injective.
